\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-29 paper.qmd
revising the binning parameters section

The performance and robustness of our model depend on four key parameters: (i) bin start position $(s_1, \ s_2)$, (ii) the total number of bins ($b$), (iii) a benchmark value used to remove low-density hexagons, and (iv) a benchmark value used to remove long edges. However, there is no analytical formula to calculate appropriate values for these parameters. The selection of these parameter values depends on the model performance computed by MSE (@sec-summary), except for the threshold used to remove long edges.

#### Bin start position and number of bins

The starting position of the hexagonal grid is important because different starting points result in different distributions of data across bins, even with the same total number of bins. This variation affects the model due to the differing number of non-empty bins. Therefore, it is necessary to evaluate various starting points with different total numbers of bins to determine which configurations are more effective at capturing the structure and fitting the model.

@@

To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$. The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$. By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-scurve (a)).

>>>>

The model fitting can be adjusted using these parameters:

- hexagon bin parameters
    - bottom left bin position $(s_1, \ s_2)$, 
    - the total number of bins ($b$), 
- bin density cutoff, to remove low-density hexagons, and 
- edge length maximum, remove long edges from \gD{} representation.

Default values are provided for each of these, but it is expected that the user will examine the MSE for a range of choices. Choosing these parameters according to MSE can be automated but it is recommended that the user examine the resulting model representation by overlaying it on the data in \pD{}. The next few subsections describe the calculation of default values, and the effect that different choices have on the model fit.

#### Hexagon bin parameters

The values $(s_1, \ s_2)$ define the position of the centroid of the bottom left hexagon. By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer sound the data. The choice of these values can have some effect on the distribution of bin counts. FIG XXX(a) illustrates this. The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown. Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.

<!--The starting position of the hexagonal grid is important because different starting points result in different distributions of data across bins, even with the same total number of bins. This variation affects the model due to the differing number of non-empty bins. Therefore, it is necessary to evaluate various starting points with different total numbers of bins to determine which configurations are more effective at capturing the structure and fitting the model.-->

@@

The default number of bins $b=b_1\times b_2$ is computed based on the sample size, by setting $b_1=n^{1/3}$, consistent with the Diaconis-Freedman rule @REF-doi:10.1007/BF01025868. The value of $b_2$ is determined analytically by $b_1, q, r_2$. Values of $b_1$ between 2 and $b_1 = \frac{n}{2}$ are allowed. FIG XXX(b) shows the effect of different choices of $b_1$ on the MSE of the fitted model.

<!-- To determine the effective $b$, candidate values are selected based on the range between the minimum and approximate maximum $b_1$, because $b_2$ is computed from $b_1$. The minimum $b_1$ is set to $2$, while the maximum number is estimated by taking the square root of $\frac{n}{2}$. By evaluating MSE across varying $b$ within this range for different $q$, helps to determine an appropriate values for $b$ and $q$ (@fig-param-scurve (a)).-->

### 2024-07-29 paper.tex
revising the binning parameters section

The performance and robustness of our model depend on four key
parameters: (i) bin start position \((s_1, \ s_2)\), (ii) the total
number of bins (\(b\)), (iii) a benchmark value used to remove
low-density hexagons, and (iv) a benchmark value used to remove long
edges. However, there is no analytical formula to calculate appropriate
values for these parameters. The selection of these parameter values
depends on the model performance computed by MSE
(Section~\ref{sec-summary}), except for the threshold used to remove
long edges.

\subsubsection{Bin start position and number of
bins}\label{bin-start-position-and-number-of-bins}

The starting position of the hexagonal grid is important because
different starting points result in different distributions of data
across bins, even with the same total number of bins. This variation
affects the model due to the differing number of non-empty bins.
Therefore, it is necessary to evaluate various starting points with
different total numbers of bins to determine which configurations are
more effective at capturing the structure and fitting the model.

@@

To determine the effective \(b\), candidate values are selected based on
the range between the minimum and approximate maximum \(b_1\), because
\(b_2\) is computed from \(b_1\). The minimum \(b_1\) is set to \(2\),
while the maximum number is estimated by taking the square root of
\(\frac{n}{2}\). By evaluating MSE across varying \(b\) within this
range for different \(q\), helps to determine an appropriate values for
\(b\) and \(q\) (Figure~\ref{fig-param-scurve} (a)).

>>>>

The model fitting can be adjusted using these parameters:

\begin{itemize}
\tightlist
\item
  hexagon bin parameters

  \begin{itemize}
  \tightlist
  \item
    bottom left bin position \((s_1, \ s_2)\),
  \item
    the total number of bins (\(b\)),
  \end{itemize}
\item
  bin density cutoff, to remove low-density hexagons, and
\item
  edge length maximum, remove long edges from \gD{} representation.
\end{itemize}

Default values are provided for each of these, but it is expected that
the user will examine the MSE for a range of choices. Choosing these
parameters according to MSE can be automated but it is recommended that
the user examine the resulting model representation by overlaying it on
the data in \pD{}. The next few subsections describe the calculation of
default values, and the effect that different choices have on the model
fit.

\subsubsection{Hexagon bin parameters}\label{hexagon-bin-parameters}

The values \((s_1, \ s_2)\) define the position of the centroid of the
bottom left hexagon. By default, this is at \(s_1 = -q, s_2 = -qr_2\),
where \(q\) is the buffer sound the data. The choice of these values can
have some effect on the distribution of bin counts. FIG XXX(a)
illustrates this. The distribution of bin counts for \(s_1\) varying
between \(-0.1-0.0\) is shown. Generally, a more uniform distribution
among these possibilities would indicate that the bins are reliably
capturing the underlying distribution of observations.

@@

The default number of bins \(b=b_1\times b_2\) is computed based on the
sample size, by setting \(b_1=n^{1/3}\), consistent with the
Diaconis-Freedman rule \citet{REF-doi:10.1007/BF01025868}. The value of
\(b_2\) is determined analytically by \(b_1, q, r_2\). Values of \(b_1\)
between 2 and \(b_1 = \frac{n}{2}\) are allowed. FIG XXX(b) shows the
effect of different choices of \(b_1\) on the MSE of the fitted model.
